MATH 245 Homework 4

### 2024-04-01 PDEs/math245_hw4.Rmd
tentative finish pde

---
title: "MATH 245 Homework 4"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-03-24"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

@@

@@

Then we plug in our formulas for $a_0$ and $a_n$:
$$ f(x) = \frac{\pi}{2} + \frac{2}{\pi} \sum_{n=1}^{\infty} \left[ \frac{1}{n^2} (-1)^{n+1} \cos{(nx)} \right]$$

@@

$$f(x)=a_0 + \sum_{n=1}^{\infty}  a_n \cos{  \left(\frac{n\pi x}{l}\right) } $$
$$f(x)=\frac{\pi^2}{3} + 4 \sum_{n=1}^{\infty} \frac{(-1)^{n}\cos{ (nx) }}{n^2}  $$

@@

The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} A_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
where 
$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
$$A_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
\begin{align*}
A_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
    &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
    &= \frac{2}{n \pi}\left[-\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)}+2\cos{\left(\frac{n\pi}{2}\right)} \right]\\
    &= \frac{2}{n \pi}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right]
\end{align*}

 $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right] \sin{(nx) }$$

@@

$$f(x)=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$
Now let $x=0$, such that $f(0)=0^2=0$ and $\cos{(nx)}=\cos{(0)}=1$.

$$0=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$
$$-\frac{\pi^2}{12}=\sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$

<!-- **(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find -->
<!-- $$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$ -->

<!-- ## Question 4 -->
<!-- Compute the complex Fourier series of the following functions: -->

<!-- **(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that -->

<!-- $$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$ -->

<!-- **(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$. -->

<!-- ## Question 5 -->

<!-- Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$. -->

<!-- $$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$ -->

>>>>

---
title: "MATH 245 Homework 4"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-01"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

@@

Since $f(x)$ and $f'(x)$ are both piecewise continuous, the Fourier series converges pointwise. However, since $f(x)$ is not continuous (only piecewise), it does not converge uniformly.

@@

Then we plug in our formulas for $a_0$ and $a_n$:
$$ f(x) = \frac{\pi}{2} + \frac{2}{\pi} \sum_{n=1}^{\infty} \left[ \frac{1}{n^2} (-1)^{n+1} \cos{(nx)} \right]$$
Since $f(x)$ and $f'(x)$ are both piecewise continuous, the Fourier series converges pointwise. Also, $f(x)$ is continuous and $f'(x)$ is piecewise continuous. Finally, the periodic boundary conditions are satisfied because $f(-\pi)=\pi=f(\pi)$. Therefore, the Fourier series also converges uniformly.

@@

$$f(x)=a_0 + \sum_{n=1}^{\infty}  a_n \cos{  \left(\frac{n\pi x}{l}\right) } $$
$$f(x)=\frac{\pi^2}{3} + 4 \sum_{n=1}^{\infty} \frac{(-1)^{n}\cos{ (nx) }}{n^2}  $$
Here, $f(x)$ and $f'(x)$ are both continuous, so the Fourier series converges piecewise. However, $f(-\pi)=\pi^2-\pi$ while $f(\pi)=\pi^2+\pi$, so the periodic boundary conditions are not satisfied and the Fourier series does not converge uniformly.

@@

The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} B_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
where 
$$B_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
$$B_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
\begin{align*}
B_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
    &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
    &= \frac{2}{n \pi}\left[-\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)}+2\cos{\left(\frac{n\pi}{2}\right)} \right]\\
    &= \frac{2}{n \pi}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right]
\end{align*}

 $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right] \sin{(nx) }$$
 $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}+2 (-1)^{n+1} \right] \sin{(nx) }$$
$$f(x)= \sum_{n_{odd}}+\sum_{n_{even}}$$
\begin{align*}
\sum_{n_{even}} &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k}\left[1+\cos{\left(\frac{2k\pi}{2}\right)}+
 2 (-1)^{2k+1} \right] \sin{(2kx) } \\
 &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k}\left[1+(-1)^n-2 \right] \sin{(2kx) } \\
 &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{(-1)^{n+1}}{2k}\sin{(2kx) }
 \end{align*}

\begin{align*}
\sum_{n_{odd}}& =
 \frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k-1}\left[1+\cos{\left(\frac{(2k-1)\pi}{2}\right)}+2 (-1)^{2k-1+1} \right] \sin{((2k-1)x) } \\
&= \frac{2}{\pi} \sum_{k=1}^{\infty} \frac{3}{2k-1}\sin{((2k-1)x) }
 \end{align*}

 $$f(x)=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{(-1)^{n+1}}{2k}\sin{(2kx) }+\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{3}{2k-1}\sin{((2k-1)x) }$$

@@

$$f(x)=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$
Parseval's Identity says that
$$ \frac{1}{2\pi}\int_{-\pi}^{\pi}(x^2)^2\;dx=a_0^2+\frac{1}{2}\sum_{n=1}^{\infty} (a_n^2+b_n^2)$$

$$ \frac{1}{2\pi}\int_{-\pi}^{\pi}(x^2)^2\;dx= \left(\frac{\pi^2}{3}\right)^2+ \frac{1}{2}\sum_{n=1}^{\infty}  \left( \frac{4(-1)^{n} }{n^2}\right)^2  $$

$$ \frac{1}{\pi}\int_{0}^{\pi}x^4\;dx= \frac{\pi^4}{9}+ \sum_{n=1}^{\infty} \frac{8}{n^4} $$
$$ \frac{1}{\pi} \frac{x^5}{5} \Bigr\rvert_{0}^{\pi}= \frac{\pi^4}{9}+ 8 \sum_{n=1}^{\infty} \frac{1 }{n^4}$$
$$  \frac{\pi^4}{5}= \frac{\pi^4}{9}+ 8 \sum_{n=1}^{\infty} \frac{1}{n^4}$$
$$ \frac{4\pi^4}{45}= 8\sum_{n=1}^{\infty} \frac{1}{n^4}$$
Thus,
$$\zeta(4)= \sum_{n=1}^{\infty} \frac{1}{n^4}=\frac{\pi^4}{90}$$

**(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find
$$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$

Take the following Fourier series and set $x=\pi$.
$$f(x)=x^2=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$

$$\pi^2=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{1 }{n^2} $$
$$\frac{2\pi^2}{3}=4 \sum_{n=1}^{\infty}  \frac{1 }{n^2} $$

Thus, 
$$\zeta(2)= \sum_{n=1}^{\infty} \frac{1}{n^2}=\frac{\pi^2}{6}$$
Now, we want to find 
$$\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} = \frac{1}{1}+\frac{1}{9}+\frac{1}{25}+\frac{1}{49}+ \;...$$
Let's also write out some terms in the series representation of $\zeta(2)$:
$$\sum_{n=1}^{\infty} \frac{1}{n^2}=\frac{1}{1}+\frac{1}{4}+\frac{1}{9}+\frac{1}{16}+\frac{1}{25}+\frac{1}{36}+\frac{1}{49}+ \;...$$
We can see that the top series is simply the bottom series without the terms where the denominator is the square of an even integer; which we can write as:
\begin{align*}
\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} 
&=\sum_{n=1}^{\infty} \frac{1}{n^2}-\sum_{n=1}^{\infty} \frac{1}{(2n)^2} \\
&=\sum_{n=1}^{\infty} \frac{1}{n^2}-\frac{1}{4}\sum_{n=1}^{\infty} \frac{1}{n^2}\\
&=\frac{\pi^2}{6}-\frac{1}{4}\frac{\pi^2}{6} \\
&=\frac{\pi^2}{8}
\end{align*}
Therefore, we conclude that $$\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} =\frac{\pi^2}{8}$$

## Question 4
Compute the complex Fourier series of the following functions:

### (a)   
Compute the complex Fourier series of $f(x) = e^x$ and show that

$$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$

The complex Fourier series of a function $f(x)$ is
$$f(x)=\sum_{n=-\infty}^{\infty} c_ne^{in\pi x / l}$$
where $$c_n= \frac{1}{2l} \int_{-l}^{l}f(x) e^{-in\pi x / l}\;dx $$

\begin{align*}
c_n&= \frac{1}{2\pi} \int_{-\pi}^{\pi}e^xe^{-in x}\;dx \\
  &= \frac{1}{2\pi} \int_{-\pi}^{\pi}e^{x(1-in)}\;dx \\
  &= \frac{1}{2\pi} \frac{1}{1-in}e^{x(1-in)}\Bigr\rvert_{-\pi}^{\pi} \\
  & = \frac{1}{2\pi} \frac{1}{1-in}\left[e^{\pi}e^{-in\pi}-e^{-\pi}e^{in\pi}\right]
\end{align*}
and since $e^{in\pi}=\cos{(n\pi)}+i\sin{(n\pi)}=(-1)^n=\cos{(-n\pi)}+i\sin{(-n\pi)}= e^{-in\pi}$,

\begin{align*}
c_n& = \frac{1}{2\pi} \frac{(-1)^n}{1-in}\left[e^{\pi}-e^{-\pi}\right]\\
&= \frac{1}{\pi} \frac{(-1)^n}{1-in} \sinh{\pi}
\end{align*}
$$f(x)=e^x=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{(-1)^n}{1-in} (\sinh{\pi})\,e^{inx}$$

Set $x=\pi$ and $x=-\pi$ and then add the two expressions, noting that $e^{in\pi}=e^{-in\pi}=(-1)^n$ and $(-1)^n(-1)^n=(-1)^{2n}=1$.
$$e^\pi=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$

$$e^{-\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$

$$\frac{e^\pi+e^{-\pi}}{2}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
But note that $\frac{e^\pi+e^{-\pi}}{2}=\cosh{\pi}$.

$$\cosh{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
Divide both sides by $\sinh{\pi}$:
$$\coth{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in}$$
Multiply the right-hand side by $\frac{1+in}{1+in}$:
$$\coth{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1+n}{1+n^2}$$

$$=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1+n^2}+\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{n}{1+n^2}$$
The second sum will disappear because the $n$ term in the numerator means that all negative terms in the series will cancel with the corresponding positive $n$ terms, while the expression evaluates to zero at $n=0$. Thus,
\begin{align*}
\coth{\pi}&=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1+n^2} \\
&=\frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} +  \;...\right)
\end{align*}
which is what we initially sought to show.

### (b)  
Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$.

The complex Fourier series of a function $f(x)$ is
$$f(x)=\sum_{n=-\infty}^{\infty} c_ne^{in\pi x / l}$$
where $$c_n= \frac{1}{2l} \int_{-l}^{l}f(x) e^{-in\pi x / l}\;dx $$

 \begin{align*}
 c_n &= \frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix}e^{-in x}\;dx \\
     &= \frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix(1-n)}\;dx
  \end{align*}

Using integration by parts,
  $$u=x, \quad du=dx,\quad dv=e^{ix(1-n)}\,dx,\quad v=\frac{1}{i(1-n)}e^{ix(1-n)} $$
$$\frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix(1-n)}\;dx = \frac{1}{2\pi} \left[\frac{x}{i(1-n)}e^{ix(1-n)}\right]^{\pi}_{-\pi}-\frac{1}{2\pi} \int_{-\pi}^{\pi}\frac{1}{i(1-n)}e^{ix(1-n)}\;dx$$
First, let's evaluate the first term on the right-hand side:
\begin{align*}
\frac{1}{2\pi} \left[\frac{x}{i(1-n)}e^{ix(1-n)}\right]^{\pi}_{-\pi} 
&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} e^{i \pi (1-n)} -\frac{-\pi}{i(1-n)} e^{-i \pi (1-n)}\right]  \\
&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} e^{i \pi}e^{-i \pi n}  -\frac{-\pi}{i(1-n)} e^{-i \pi} e^{i \pi n} \right]  \\
&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} (-1)(-1)^n  -\frac{-\pi}{i(1-n)} (-1)(-1)^n \right] \\
&= \frac{1}{2\pi} \left[ \frac{2\pi}{i(1-n)} (-1)^{n+1}  \right] \\
&= \frac{(-1)^{n+1}} {i(1-n)} 
\end{align*}

Now we turn to the integral on the right-hand side:
\begin{align*}
-\frac{1}{2\pi} \int_{-\pi}^{\pi}\frac{1}{i(1-n)}e^{ix(1-n)}\;dx &= \frac{1}{2\pi} \left[\frac{e^{ix(1-n)}}{(1-n)^2}\right]^{\pi}_{-\pi}  \\
&= \frac{1}{2\pi} \left[ \frac{e^{i\pi (1-n)}}{(1-n)^2}  -\frac{e^{-i \pi (1-n)}}{(1-n)^2}\right] \\
&= \frac{1}{2\pi} \left[ \frac{(-1)^{n+1}}{(1-n)^2}  -\frac{(-1)^{n+1}}{(1-n)^2}\right] \\
&=0 
\end{align*}

Thus, $c_n=\frac{(-1)^{n+1}}{i(1-n)}$. Plugging this into the equation for a complex Fourier series, we find that

$$xe^{ix}=\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx}$$

Now, we use Euler's identity
$$e^{ix}=\cos{x}+i\sin{x}$$
$$\cos{x}=\frac{e^{ix}+e^{-ix}}{2}$$
\begin{align*}
x \cos{x} &=  \frac{x\left( e^{ix}+ e^{-ix} \right) } {2} \\
          &= \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx} + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{-inx}\\
          &= \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(-nx)}+i\sin{(-nx)})  \\
          &=\frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(nx)}-i\sin{(nx)}) 
\end{align*}

The sin terms cancel and the cos terms add, yielding
$$x \cos{x} =x \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} \cos{nx} $$
Similarly,

\begin{align*}
\sin{x} &=\frac{e^{ix}-e^{-ix}}{2i} \\
        &= \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx} - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{-inx} \\
        &= \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(-nx)}+i\sin{(-nx)})  \\
        &=\frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(nx)}-i\sin{(nx)})
\end{align*}

The $\cos$ terms cancel and the $\sin$ terms add, yielding
$$\frac{x}{i} \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(- \sin{nx}) $$
Multiplying the two $i$s together cancels the negative sign, meaning the Fourier series becomes:
$$x \sin{x} =x \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {(1-n)}\sin{nx} $$

## Question 5

Find the function represented by the new series which is obtained by term-wise integration of the following series from $0$ to $x$.

$$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$

$$ \int_{0}^{x} \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$

$$\left[ \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2} \right]\Bigr\rvert_{0}^{x}= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$
Therefore,
$$\int_{0}^{x} \log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$
